Symmetric Peer-to-Peer Applications

### 2022-10-24 tpd-22.tex
...

\section{Evaluation}
\label{sec.eval}

In this section, we evaluate if (and to what extent) our design meets its goals
to ensure that all peers (a) meet event deadlines, (b) advance in sync in real
time, and (c) can leave and join the network and remain symmetric.
%
We perform simulation varying the network latency and peer churn, as well as
the frequency and deadlines of events.
%
We then measure (a) the recurrence of time travels, (b) the real-time pace of
peers, and (c) the recoverability from churns, such that they validate our
goals above, respectively.

Regarding (a), if a peer experiences a time travel, it means that it received
an event after its deadline.
We then count all time travel occurrences in the evaluation.
%
Regarding (b), we use an absolute simulation clock to track, in real time, the
execution pace of each peer.
% in comparison to the reference peer, which is the one most advanced in time.
We then count the mismatches between the clocks in the evaluation.
%
Regarding (c), we make peers leave and join the network periodically, possibly
creating partitions.
We then assert that they all catch up in real time within a short period in the
evaluation.

@@

We perform simulations with 21 peers using the mixed topology of
Figure~\ref{fig.topo}, with peers organized in straight lines and cycles.
The topology has an average of 5 hops between peers (e.g., \code{4-16} are 7
hops away).
%
In order to have an absolute clock, we use a single machine to simulate all
peers, allowing us to merge their logs into a single file with a comparable
shared timeline.
%
We use \emph{NetEm}~\cite{netem} to simulate the network latency with a normal
distribution.
% (e.g., \code{tc qdisc add dev lo root netem delay 200ms 40ms distribution normal}).
%
We vary
    (i)   the network latency (\code{5-500ms} between each peer),
    (ii)  the rate of events (\code{5-200 evt/sec} in the whole network), and
    (iii) the delay multiplier (\code{1-5x} considering the latency and the
          average hops between peers).
%
Figure~\ref{fig.mult} shows the event delays considering the network latency
and chosen multiplier.
For instance, a \code{50ms} latency with a \code{2x} multiplier results in
\code{500ms} event delays, i.e., each event is triggered with this delay to
compensate the given latency and fixed average of 5 hops.
%
These variations result in $108$ combinations.
We executed each combination $3$ times for $5$ minutes each.
We also make some modifications to evaluate peer churn, which requires to
re-execute the simulation.
The experiments take around $40$ hours to complete.

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{baks}
  \caption{
    \label{fig.baks}
The \code{5x}, \code{2x}, and \code{1x} delay regions each contains $36$
measures for multiple configurations of event rate (\code{5-200evt/min} and
network latency (\code{5-500ms}).
%
The color thresholds (\emph{white}, \emph{yellow}, \emph{red}, \emph{black})
are arbitrary but help to distinguish between regions of interest.
%
The experiments uses $1$ process for each peer in a single Linux machine
(\emph{i7/16GB}).
    }
\end{figure*}

We now detail the results and how we instrument the middleware implementation
to measure the properties above:
%
\begin{enumerate}
%
\item \textbf{Recurrence of time travels:}
Every time a peer is ahead of time and needs to travel back, we output its id
and how much ticks it needs to travel back.
At the end, we measure the percentage of the sum of all time travel ticks over
the total simulation ticks.
As an example, if all peers sum $1000$ time travel ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
Figure~\ref{fig.baks} presents the results considering all combinations of
parameters (i), (ii), and (iii) above.
The colors indicate the feasibility of each tested scenario
    (\emph{white=good}, \emph{yellow=moderate}, \emph{red=poor},
    \emph{black=faulty}).
%
For instance, the three framed rectangles ($0.1$, $3.1$, and $7.2$) represent
the scenarios with $50ms$ network delay, generating $25evt/min$.
Each rectangle uses a different delay multiplier, showing a good performance
for $5x$, moderate for $2x$, and poor for $1x$.
%
\item \textbf{Real-time pace of peers:}
All peers output their current ticks continuously, allowing us to compare them
in the shared timeline.
Every time we see a greater tick never seen in the timeline before, we count
smaller ticks from other peers, which constitute time violations.
As an example, if all peers sum $1000$ of such smaller ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
We also start each peer with a random delay of up to $10s$ to force a
synchronization mismatch at the beginning.
%
Considering all scenarios, the average is negligibly under $0.2\%$, with no
significant variance.
Therefore, we omit the resulting table akin to Figure~\ref{fig.baks}), which
would not provide further insights.
%
\item \textbf{Correctness of unstable peers:}
We tweak our simulation to make each peer remain $40s$ online and $20s$
offline in the average, including peers \code{9-11} in the middle of the
network.
This creates partitions in the network, which generate events out of sync that
require constant resynchronization between peers.
%
Then, to evaluate the network correctness under such high churn, we count how
much time the recovering peers take to catch up in real time:
When a peer becomes online, we take the current maximum tick considering all
reachable peers, and then count the time the recovering peer takes to catch up.
We also assert that they all reach the same final state eventually.
%
Considering all scenarios, the average is $1.3$ seconds of recovering time,
with no significant variance.
Therefore, we also omit the resulting table akin to Figure~\ref{fig.baks}).
%
\end{enumerate}

@@

Our results show that goals (b) and (c) are met in all scenarios consistently,
and we consider them nearly optimal:
Regardless of the network latency, event rate, and delay multiplier,
    the variation in the pace of peers is under $0.2\%$ (goal b), and
    recovering peers take $1.3$ seconds to catch up in time (goal c).
Regarding goal (c), note that the middleware is configured to make small steps
towards smooth time transitions, which take $1$ second to complete.
Therefore, the average of $1.3$ seconds shows that peers need no more than 2
time travels while recovering.

Considering goal (a), we need to take a closer look at Figure~\ref{fig.baks}.
The color patterns indicate clear boundaries for the scenarios that can meet
event deadlines with the desired performance.
%
For instance, high event delays (column $5x$) make all scenarios viable,
even with high latency and event frequency.
As an application example, chats are not affected by high event delays.
Nevertheless, the higher are the event delays, the less applications suit the
middleware.
%
For the lowest possible delay (column $1x$), we see that the middleware can
either handle low latency with high event frequency (e.g., $5ms$ with
$100evt/min$) or high lantency with low event frequency (e.g., $100ms$ with
$5evt/min$).

\section{TODO}

TODO: simpler than vector clocks, absolute times are correct

- p2p links por fora, como um servico TCP local

We make some assumptions about the network, which are assigned externally to
the middleware:
%
\begin{itemize}
\item Peers identifiers are assigned externally to the middleware.
\item Topology is dynamic, but also assigned externally

\end{itemize}
%
Except for the last item, we claim that the others can be solved externally,
without affecting the middleware.
For instance,
- orthogonal and would not affect the other algorithms in the middleware

, which we claim that could be
solved further:
- ids are outside
- topology, dynamic but outside
- fixed maximum nodes
- fixed latency
We claim that Use external protocols
For instance, latency is only a constant, no problem being a variable

- non-malicious nodes

\section{Application}
\label{sec.app}

>>>>

%\newpage
\section{Evaluation}
\label{sec.eval}

In this section, we evaluate if our design meets the goals to ensure that all
peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.
%
We perform simulations varying the network latency and peer churn, as well as
the frequency and deadlines of events.
%
We then measure
    (a) the recurrence of time travels,
    (b) the real-time pace of peers, and
    (c) the recoverability from churns,
such that they validate our goals above, respectively.

Regarding (a), if a peer experiences a time travel, it means that it received
an event after its deadline.
We then count all time travel occurrences in our evaluation.
%
Regarding (b), we use an absolute simulation clock to track, in real time, the
execution pace of each peer.
% in comparison to the reference peer, which is the one most advanced in time.
We then count the mismatches between the clocks in our evaluation.
%
Regarding (c), we make peers leave and join the network periodically, possibly
creating partitions.
We then assert that they all catch up in real time within a short period in our
evaluation.

@@

We perform simulations with 21 peers using the mixed topology of
Figure~\ref{fig.topo}, with peers organized in straight lines and cycles.
The topology has an average of 5 hops between peers (e.g., \code{4-16} are 7
hops away).
%
We run the application main loop at $50 FPS$ ($20ms$ per frame) and calculate
the middleware overhead at each iteration, which is $9us$ in the average,
corresponding to less than $1/2000$ of the CPU time.
%
In order to have an absolute clock, we use a single machine to simulate all
peers, allowing us to merge their logs into a single file with a comparable
shared timeline.
%
We use \emph{NetEm}~\cite{netem} to simulate the network latency with a normal
distribution.
% (e.g., \code{tc qdisc add dev lo root netem delay 200ms 40ms distribution normal}).
%
We vary
    (i)   the network latency ($5-500ms$ between each peer),
    (ii)  the rate of events ($5-200~evt/sec$), and
    (iii) the delay multiplier ($1-5x$).
%
Figure~\ref{fig.mult} shows the event delays considering the network latency
and chosen multiplier.
For instance, a $50ms$ latency with a $2x$ multiplier results in $500ms$ event
delays, i.e., each event is triggered with this delay to compensate the given
latency and fixed average of 5 hops.
%
These variations result in $108$ combinations.
We executed each combination $3$ times for $5$ minutes each.
We also make some modifications to evaluate peer churn, which requires to
re-execute the simulation.
The experiments take around $40$ hours to complete.

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{baks}
  \caption{
    \label{fig.baks}
The $5x$, $2x$, and $x$ delay regions each contains $36$ measures for multiple
configurations of event rate ($5-200~evt/min$ and network latency ($5-500ms$).
%
The color thresholds (\emph{white}, \emph{yellow}, \emph{red}, \emph{black})
are arbitrary but help to distinguish between regions of interest.
%
The experiments uses $1$ process for each peer in a single Linux machine
(\emph{i7/16GB}).
    }
\end{figure*}

We now detail the results and how we instrument the middleware implementation
to measure the properties above:
%
\begin{enumerate}
%
\item \textbf{Recurrence of time travels:}
Every time a peer is ahead in time and needs to travel back, we output its id
and how much ticks it needs to travel back.
At the end, we measure the percentage of the sum of all time travel ticks over
the total simulation ticks.
As an example, if all peers sum $1000$ time travel ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
Figure~\ref{fig.baks} presents the results considering all combinations of
parameters (i), (ii), and (iii) above.
The colors indicate the feasibility of each tested scenario
    (\emph{white=good}, \emph{yellow=moderate}, \emph{red=poor},
    \emph{black=faulty}).
%
For instance, the three framed rectangles ($0.1$, $3.1$, and $7.2$) in the
figure represent the scenarios with $50ms$ network delay, generating
$25~evt/min$.
Each rectangle uses a different delay multiplier, showing a good performance
for $5x$, moderate for $2x$, and poor for $1x$.
%
\item \textbf{Real-time pace of peers:}
All peers output their current ticks continuously, allowing us to compare them
in the shared timeline.
Every time we see a greater tick never seen in the timeline before, we count
smaller ticks from other peers, which constitute time violations.
As an example, if all peers sum $1000$ of such smaller ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
We also start each peer with a random delay of up to $10s$ to force a
synchronization mismatch at the beginning.
%
Considering all scenarios, the average is negligibly under $0.2\%$, with no
significant variance.
Therefore, we omit the resulting table akin to Figure~\ref{fig.baks}), which
would not provide further insights.
%
\item \textbf{Correctness of unstable peers:}
We tweak our simulation to make each peer remain $40s$ online and $20s$
offline in the average, including peers \code{9-11} in the middle of the
network.
This creates partitions in the network, which generate events out of sync that
require constant resynchronization between peers.
%
Then, to evaluate the network correctness under such high churn, we count how
much time a recovering peer takes to catch up in real time:
When a peer becomes online, we take the current maximum tick considering all
reachable peers, and then count the time the recovering peer takes to catch up.
We also assert that they all reach the same final state eventually.
%
Considering all scenarios, the average is $1.3$ seconds of recovering time,
with no significant variance.
Therefore, we also omit the resulting table (akin to Figure~\ref{fig.baks}).
%
\end{enumerate}

@@

Our results show that goals (b) and (c) are met in all scenarios consistently,
and we consider them nearly optimal:
Regardless of the network latency, event rate, and delay multiplier,
    the variation in the pace of peers is under $0.2\%$ (goal b), while
    recovering peers take $1.3$ seconds to catch up in time (goal c).
Also regarding goal (c), note that the middleware makes smooth time transitions
in small steps that take $1$ second to complete.
Therefore, the average of $1.3$ seconds shows that peers need no more than 2
time travels while recovering.

Considering goal (a), we need to take a closer look at Figure~\ref{fig.baks}.
The color patterns indicate clear boundaries for the scenarios that can meet
event deadlines with the desired performance.
%
For instance, high event delays (column $5x$) make all scenarios viable,
even with high latency and event frequency.
As an application example, chats are not affected by high event delays.
Nevertheless, the higher are the event delays, the less applications suit the
middleware.
%
For the lowest possible delay (column $1x$), we see that the middleware can
either handle low latency with high event frequency (e.g., $5ms$ with
$100evt/min$) or high lantency with low event frequency (e.g., $100ms$ with
$5evt/min$).
%
The color patterns in the table provide the necessary information to confront
an application with its feasibility or desired performance.

%\section{Application}
%\label{sec.app}
